Data Munging

### 2019-12-19 scripts/dataMunge.Rmd
Estimated missing max buouancy frequency data

# there are three non-finite bf values, lets have a look
bf.na <- bf %>% filter(!is.finite(BF)) %>%
  select(SITE_ID, VISIT_NO)

@@

# ONe of the three values comes from a site with other non-zero measurements.  Probably OK.
# two sites have only a single NA for BF becasue they only have a single temp measurement, 
# presumably these are shallow sites, therefore unlikely to be stratified.  Assign a bf
# very close to 0.
bf %>% filter(SITE_ID %in% bf.na$SITE_ID) 
temp2 %>% filter(SITE_ID %in% bf.na$SITE_ID)

@@

Extract the max buoyancy frequency value for each waterbody. 
``` {r}
# Finally, extract max buoyancy for each SITE_ID x VISIT_NO
bf.max <- bf %>% group_by(SITE_ID, VISIT_NO) %>%
  summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
  ungroup()
```

`r bf.max %>% summarize(n.na = sum(is.na(bf.max$MAX.BF))) %>% pull()` of the profiles do not have a maximum buoyancy frequency value.  Very good.

>>>>

# there are three non-finite bf values, lets have a look
bf.na <- bf %>% filter(!is.finite(BF)) %>%
  select(SITE_ID, VISIT_NO) %>%
  mutate(SITE.VISIT = paste0(SITE_ID, VISIT_NO))

@@

# These site have a single temp measurement made at shallow depths (<1m).  Presumably these are 
# shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0.
bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
temp2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)

@@

Extract the max buoyancy frequency value for each waterbody after converting negative BF values to 0, per Webb.
``` {r}
# Finally, extract max buoyancy for each SITE_ID x VISIT_NO
bf.max <- bf %>% 
  mutate(BF = ifelse(BF < 0,
                     0,
                     BF)) %>%
  group_by(SITE_ID, VISIT_NO) %>%
  summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
  ungroup()
```

`r bf.max %>% summarize(n.na = sum(!is.finite(bf.max$MAX.BF))) %>% pull()` of the site have a non-finite value for maximum buoyancy frequency value.  These are the three shallow sites with only a single temperature measurement.  Assign these the minimum bf.max value.
``` {r}
bf.max %>% filter(!is.finite(MAX.BF))

min.max.bf <- bf.max %>% 
  filter(is.finite(MAX.BF)) %>%
  summarize(min.max.bf = min(MAX.BF))

bf.max <- bf.max %>%
  mutate(MAX.BF = replace(MAX.BF, 
                          !is.finite(MAX.BF), # if not finite
                          min.max.bf$min.max.bf)) # then min

# that took care of it
bf.max %>% filter(!is.finite(MAX.BF))
```
